Lensing in the Schwarzschild Spacetime Applicable to Neutron Stars

### 2023-03-28 src/tex/ms.tex
expanded inegrals

It will be convenient to transform variables from $u$ to $w = u/\rNS$, whence 
\begin{equation}
    \chi^\phi\left( \rNS \right) = \int_{1}^{\infty} \dd w \, \frac{b}{\rNS w^2 \sqrt{1 - \frac{b^2\left(1 - \frac{2}{\rNS w} \right)}{\rNS^2 w^2}}}.
\end{equation}
It is also convenient to make the substitution 
\begin{equation}
    b = x \bNS = x \rNS \left( 1 + \frac{1}{\rNS} + \frac{3}{2 \rNS^2} + \ldots \right),
\end{equation}
whence $-1 < x < 1$ for rays that impact the neutron star surface. This produces 
\begin{equation}
    \chi^\phi\left( \rNS \right) = \int_1^\infty \dd w \, \frac{x \left( 1 + \frac{1}{\rNS} + \ldots \right)}{w^2 \sqrt{1 - \frac{x^2 \left( 1 + \frac{2}{\rNS} + \ldots \right)\left( 1 - \frac{2}{\rNS w} \right)}{w^2}}}.
\end{equation}

@@

Let's calculate $\frac{d\chi^t(\lambda)}{d\lambda}$ and $\frac{d\chi^r(\lambda)}{d\lambda}$ assuming $\theta=\pi/2$ and therefore $p^\theta = 0$. The we have
in null geodesic:
%
\begin{linenomath}\begin{align}
    \label{eq:nullgeo}
    -(1-\frac{2}{r})(p^t)^2 + \frac{1}{1-\frac{2}{r}}(p^r)^2 + r^2\sin^2 \theta (p^\phi)^2 = 0
\end{align}\end{linenomath}
%
Solving eq. \ref{eq:nullgeo} leads to $p^t = \frac{d\chi^t(\lambda)}{d\lambda} = \frac{e}{1-\frac{2}{r}}$ and 
%
\begin{linenomath}\begin{align}
    \label{eq:pr-squared}
    \left(\frac{d\chi^r(\lambda)}{d\lambda}\right)^2 = \left(\frac{e^2}{1-\frac{2}{r}} - \frac{l}{r^2}\right) \left(1-\frac{2}{r}\right)
\end{align}\end{linenomath}
%
which, in terms of $b$ becomes
%
\begin{linenomath}\begin{align}
    \label{eq:pr-b}
    \frac{d\chi^r(\lambda)}{d\lambda} = e^2\sqrt{\left(\frac{1}{1-\frac{2}{r}} - \frac{b}{r^2}\right) \left(1-\frac{2}{r}\right)}
\end{align}\end{linenomath}
%

\bibliography{bib}

>>>>

It will be convenient to transform variables from $u$ to $w = u/\rNS$, whence 
\begin{equation}
    \label{eq:w-substitution}
    \chi^\phi\left( \rNS \right) = \int_{1}^{\infty} \dd w \, \frac{b}{\rNS w^2 \sqrt{1 - \frac{b^2\left(1 - \frac{2}{\rNS w} \right)}{\rNS^2 w^2}}}.
\end{equation}
It is also convenient to make the substitution 
\begin{equation}
    b = x \bNS = x \rNS \left( 1 + \frac{1}{\rNS} + \frac{3}{2 \rNS^2} + \ldots \right),
\end{equation}
whence $-1 < x < 1$ for rays that impact the neutron star surface. This produces 
\begin{equation}
    \chi^\phi\left( \rNS \right) = \int_1^\infty \dd w \, \frac{x \left( 1 + \frac{1}{\rNS} + \ldots \right)}{w^2 \sqrt{1 - \frac{x^2 \left( 1 + \frac{2}{\rNS} + \ldots \right)\left( 1 - \frac{2}{\rNS w} \right)}{w^2}}}.
\end{equation}

@@

Let's calculate $\frac{d\chi^t(\lambda)}{d\lambda}$ and $\frac{d\chi^r(\lambda)}{d\lambda}$ assuming $\theta=\pi/2$ and therefore $p^\theta = 0$. The we have
in null geodesic:
%
\begin{linenomath}\begin{align}
    \label{eq:nullgeo}
    -(1-\frac{2}{r})(p^t)^2 + \frac{1}{1-\frac{2}{r}}(p^r)^2 + r^2\sin^2 \theta (p^\phi)^2 = 0
\end{align}\end{linenomath}
%
Solving eq. \ref{eq:nullgeo} leads to $p^t = \frac{d\chi^t(\lambda)}{d\lambda} = \frac{e}{1-\frac{2}{r}}$ and 
%
\begin{linenomath}\begin{align}
    \label{eq:pr-squared}
    \left(\frac{d\chi^r(\lambda)}{d\lambda}\right)^2 = \left(\frac{e^2}{1-\frac{2}{r}} - \frac{l}{r^2}\right) \left(1-\frac{2}{r}\right)
\end{align}\end{linenomath}
%
which, in terms of $b$ becomes
%
\begin{linenomath}\begin{align}
    \label{eq:pr-b}
    \frac{d\chi^r(\lambda)}{d\lambda} = e\sqrt{\left(\frac{1}{1-\frac{2}{r}} - \frac{b}{r^2}\right) \left(1-\frac{2}{r}\right)}
\end{align}\end{linenomath}
%
Therefore, the integral \ref{eq:timedelay} becomes:
%
\begin{linenomath}\begin{align}
    \label{eq:timedelayfull}
    \Delta t = \int_{r_{NS}}^{\infty} dr \left[\left(\frac{1}{1-\frac{2}{r}}\right) - \frac{1}{\sqrt{\left(1-\frac{2}{r}\right)^2 - \frac{b}{r^2}\left(1-\frac{2}{r}\right)^3}}\right]
\end{align}\end{linenomath}
%
Following the substitution in Eq. \ref{eq:w-substitution}, we then get
%
\begin{linenomath}\begin{align}
    \label{eq:tdelay-with-w}
    \Delta t = \int_{1}^{\infty} \frac{dw}{r_{NS}} \left[\left(\frac{1}{1-\frac{2}{wr_{NS}}}\right) - \frac{1}{\sqrt{\left(1-\frac{2}{wr_{NS}}\right)^2 - \frac{b}{(wr_{NS})^2}\left(1-\frac{2}{wr_{NS}}\right)^3}}\right]
\end{align}\end{linenomath}
%
It is also convenient to make the substitution 
\begin{equation}
    b = x \bNS = x \rNS \left( 1 + \frac{1}{\rNS} + \frac{3}{2 \rNS^2} + \ldots \right),
\end{equation}
whence $-1 < x < 1$ for rays that impact the neutron star surface. This produces 
%
\begin{linenomath}\begin{align}
    \label{eq:tdelay-with-x}
    \Delta t = \int_{1}^{\infty} \frac{dw}{r_{NS}} \left[\left(\frac{1}{1-\frac{2}{wr_{NS}}}\right) - \frac{1}{\sqrt{\left(1-\frac{2}{wr_{NS}}\right)^2 - \frac{x \rNS \left( 1 + \frac{1}{\rNS} + \frac{3}{2 \rNS^2} + \ldots \right)}{(wr_{NS})^2}\left(1-\frac{2}{wr_{NS}}\right)^3}}\right]
\end{align}\end{linenomath}
%
\bibliography{bib}
